---
title: "R Notebook"
output: html_notebook
---
```{r setup}
pacman::p_load(marmap, mapdata, RColorBrewer, classInt, ggplot2, install = FALSE)
# library(marmap)
# library(mapdata)
# library(RColorBrewer)
# library(classInt)
# library(ggplot2)
```
# Get data
- These numbers are the western longitude, the eastern longitude, the southern latitude and the northern latitude.  
- getNOAA.bathy is from the marmap package.
```{r}
dat <- getNOAA.bathy(-78,-64,37,43,res=1, keep=TRUE)

data(nw.atlantic)
atl <- as.bathy(nw.atlantic)
```

# Create color palettes
```{r}
blues <- c("lightsteelblue4", "lightsteelblue3", "lightsteelblue2", "lightsteelblue1")
greys <- c(grey(0.6), grey(0.93), grey(0.99))
```
# Plot data
```{r}
plot(dat
     # set x and y axis limits
     , xlim=c(-76.5,-65.5)
     , ylim=c(38,41.5)
     # im plots the terrain image
     ,im=TRUE
     # plots land as grey and sea as blue?  looks like it changes where the land sea boundary is
     , land=TRUE
     # add in color palettes defined above - defines blues as sea and greys as land, c(low point, high point, color) where in this line, the lows are darkest, with 0 (beach) being lightest and for land the lows are lightest and mountains are lightest.
     , bpal=list(c(min(dat),0,blues),c(0,max(dat),greys))
     # lwd controls the alpha of the contour lines, but not the bathymetric contour lines, it looks like there are 2 sets here.
     , lwd=.09
     # las = include more axis labels?
     , las=1
     # percent size of the axis titles
     # ,cex.lab=1.75
     # percent size of the tick labels
     # , cex.axis=1.75
     # no apparent what this one does but it must be size somehow - probably main title which is lacking here
     # , cex.main=1.75
     # no apparent what this one does but it must be size somehow - probably subtitle which is lacking here
     # , cex.sub=1.75 
     )

# this plots the border between land and sea
# map("worldHires", res=0, lwd=0.7, add=TRUE)
# these plot the contour lines
plot(dat, deep=-25, shallow=-25, step=0, lwd=0.1, drawlabel=TRUE, add=TRUE)
plot(dat, deep=-50, shallow=-50, step=0, lwd=0.1, drawlabel=TRUE, add=TRUE)
plot(dat, deep=-75, shallow=-75, step=0, lwd=0.1, drawlabel=TRUE, add=TRUE)
plot(dat, deep=-100, shallow=-100, step=0, lwd=0.1, drawlabel=TRUE, add=TRUE)

```
Can we ggplot this? [Directions](https://rdrr.io/cran/marmap/man/autoplot.bathy.html)

- plot the original contours and map borders  

```{r}
autoplot(dat)
```
- plot the tile - color shade of depth  

```{r}
autoplot(dat, geom = c("tile"))
```

- plot both!
```{r}
autoplot(dat, geom=c("raster", "contour"))

  # geom names can be abbreviated
autoplot(dat, geom=c("r", "c"))
```

- do not highlight the coastline
```{r}
autoplot(dat, coast=FALSE)
```

    
- better colour scale
```{r}
autoplot(dat, geom=c("r", "c")) +
  scale_fill_gradient2(low="dodgerblue4", mid="gainsboro", high="darkgreen")
```


- set aesthetics - even though it gives and "unknown", it changes the contour lines to a 0.1px white line
```{r}
autoplot(dat, geom=c("r", "c"), colour="white", size=0.1)
```
	
  
- topographical colour scale, see ?scale_fill_etopo
```{r}
autoplot(dat, geom=c("r", "c"), colour="white", size=0.1) + 
  scale_fill_etopo()

autoplot(atl, geom=c("r", "c"), colour="white", size=0.1) + 
  scale_fill_etopo()
```
	
- add sampling locations
```{r}
data(metallo)

last_plot() + geom_point(aes(x=lon, y=lat), data=metallo, alpha=0.5)
```

- an alternative contour map making use of additional mappings
- see ?stat_contour in ggplot2 to understand the ..level.. argument
```{r}
autoplot(dat, geom="contour", mapping=aes(colour=..level..))
```


